Inference of proto-neutron star properties from gravitational-wave data in core-collapse supernovae.

### 2020-10-30 introduction.tex
Finished my revision of the introduction

 % {We base this work in the PNS mode analysis performed by \citep{Torres:2018, Torres:2019a}, which explored
\mab{The eigenmode spectrum of the region within the shock (including the PNS and the post-shock region) 
using results from 2D CCSN numerical simulations as a background studied in \citep{Torres:2018, Torres:2019a}
shows a good match to the mode frequencies computed and the features observed in the GW spectrum of the same
simulation (specially when space-time perturbations are included \citep{Torres:2019a}).}
This result reveals that it is posible to perform CCSN asteroseismology {under realistic conditions} and serves as a starting point to carry out inference of astrophysical parameters of PNSs. Authors in {\citep{Torres:2019b} went one step further showing that it was possible to derive simple relations} between the instantaneous frequency of the g-mode and the mass and radius of the PNS {at each time of the evolution}. These relations are universal in the sense that they do not depend {on the equation of state (EOS) used} or the mass of the progenitor, {and only weakly on} the numerical code used {(see discussion in section~\ref{sec:simulations})}. {Similar relations have been found by \citep{Sotani:2020a,Sotani:2020b}, which also found that the universal relations do not depend on the dimensionality (1D, 2D or 3D) of the numerical simulation used as a background.}

{In this work, we present a method to infer from the GW data alone, the time evolution of some properties of the PNS, namely a combination of its mass and radius. For this purpose we have developed an algorithm to  
extract the time-frequency evolution of the main feature in the spectrograms of the GW emission of 2D simulations of CCSN. This feature corresponds to the $^2\rm{g}_2$ mode, according to the nomenclature used in \citep{Torres:2019b} (different authors may have slightly different naming convention). Next, we use the universal relations obtained by \citep{Torres:2019b}{, based on a set of 1D simulations,} to infer the time evolution of the ratio $M_{\rm PNS}/R_{\rm PNS}^2$, being 
$M_{\rm PNS}$ and $R_{\rm PNS}$ the mass and radius of the PNS.} Using 2D CCSN waveform corresponding to different progenitor masses we estimate the performance of the algorithm for current and future generation of ground-based GW detectors.

This paper is organised as follows. Section II describes the details of the CCSN simulations used in the paper. Section III focuses on the algorithm that extracts the time evolution of a combination of the mass and radius of the PNS corresponding to a g-mode. Section IV shows the performance of the data analysis method with simulated GW detectors data. Finally, we discuss the results in section V.

>>>>

Using results from 2D CCSN numerical simulations as a background with respect to which to solve an eigenvalue problem~\citep{Torres:2018, Torres:2019a} found that the eigenmode spectrum of the region within the shock (including the PNS and the post-shock region) 
shows a good match to the mode frequencies and to the features observed in the GW spectrum of the same simulations (specially when space-time perturbations are included \citep{Torres:2019a}). \tf{This last sentence might be improved.}
This reveals that it is posible to perform CCSN asteroseismology {under realistic conditions} and serves as a starting point to carry out inference of astrophysical parameters of PNSs. 
Further work was presented in {\citep{Torres:2019b} who found that it is possible to derive simple relations} between the instantaneous frequency of the g-mode and the mass and radius of the PNS {at each time of the numerical evolutions}. These relations are universal as they do not depend {on the equation of state (EOS) or on the mass of the progenitor {and they only depend weakly on} the numerical code used {(see discussion in Section~\ref{sec:simulations})}. {Similar universal relations have been discussed by \citep{Sotani:2020a,Sotani:2020b} who also found that they do not depend on the dimensionality (1D, 2D or 3D) of the numerical simulation used as a background.

{In this work we introduce a method to infer PNS properties, namely a combination of the mass and radius, using GW information. For this purpose we have developed an algorithm to  
extract the time-frequency evolution of the main feature in the spectrograms of the GW emission of 2D simulations of CCSN. This feature corresponds to the $^2\rm{g}_2$ mode, according to the nomenclature used in \citep{Torres:2019b} (different authors may have slightly different naming convention). Next, we use the universal relations obtained by \citep{Torres:2019b}{, based on a set of 1D simulations,} to infer the time evolution of the ratio $M_{\rm PNS}/R_{\rm PNS}^2$ (the PNS surface gravity), where  $M_{\rm PNS}$ and $R_{\rm PNS}$ are the mass and the radius of the PNS.} Finally, using 2D CCSN waveform corresponding to different progenitor masses we estimate the performance of the algorithm for current and future generation of ground-based GW detectors.

This paper is organized as follows. Section II describes the details of the CCSN simulations used in the paper. The algorithm that we employ to extract the time evolution of the PNS surface gravity is discussed in Section III. Section IV shows the performance of our inference method with simulated GW detectors data and presents our main results. Our findings are summarized in section V. Appendix A discusses specific details related to the reconstruction of the g-mode.
